### 2012-11-29 inst/examples/comparing-methods.Rmd
basic example comparing methods

```{r}
require(MASS)
require(ggplot2)
require(kernlab)
opts_knit$set(upload.fun = socialR::notebook.url)
```

@@

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
                  y = c(-2,  0,  1,  2, -1, -1))
#x <- -5:5
#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
l <- 1
sigma_n <- 0.
```

Radial basis function/Gaussian kernel:

```{r}
  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

@@

```{r}
  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  L <- chol(K + sigma_n^2 * I)
  alpha <- solve(t(L), solve(L, obs$y))
  k_star <- cov(obs$x, X)
  Y <- t(k_star) %*% alpha
  v <- solve(L, k_star)
  Var <- cov(X,X) - t(v) %*% v
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
```

@@

``` {r}
  cov_xx_inv <- solve(K + sigma_n^2*I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
```

@@

``` {r}
ef <- numeric(length(X))
cf <- matrix(0, nrow=length(X), ncol=length(X))

A <- as.numeric( cov(obs$x[1], obs$x[1]) )
mu <- obs$y[1]

for(i in 2:length(obs$x)){
  mu <- obs$y[i] + cov(obs$x[i], obs$x[1:(i-1)]) %*% (obs$y[1:(i-1)] - mu) / (A + sigma_n^2)
  A <- as.numeric( cov(obs$x[i], obs$x[i]) - cov(obs$x[i], obs$x[1:(i-1)]) %*% cov(obs$x[1:(i-1)], obs$x[i]) / (A + sigma_n^2) )
}

ef <- cov(X, obs$x) %*% (obs$y - mu) / A
cf <- cov(X,X) - cov(X, obs$x) %*% cov(obs$x, X) / A

@@

``` {r}
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
y_p <- predict(gp, X)
```

These are the same as `y_p`, explicitly showing the $E_y = K(x, x_i) \alpha$ multiplication

``` {r}
y_mult <-  kernelMult(kernelf(gp), X, xmatrix(gp), as.matrix(alpha(gp))) 
y_se <- sapply(X, function(x) cov(x, obs$x) %*% alpha(gp))
rbind(as.numeric(y_p), as.numeric(y_mult), as.numeric(y_se))[,1:10]
```

Which demonstrates the kernel trick `alpha = cov_xx_inv %*% obs$y`

### mlegp method

``` {r}
require(mlegp)
out <- mlegp(obs$x, obs$y, nugget = 1)
y_m <- predict(out, as.matrix(X))
```

@@

Compare alpha from Cholesky and alpha from `kernlab` to $K(x,x)^{-1} \vec y$

```{r}
cov_xx_inv %*% obs$y
alpha
alpha(gp)
```

>>>>

```{r}
require(MASS)
require(ggplot2)
require(kernlab)
require(nonparametricbayes)
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
```

@@

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
                  y = c(-2,  0,  1,  2, -1, -1))
l <- 1
sigma_n <- 0.1
```

```{r results="hide", eval=FALSE, include=FALSE}
#knit("beverton_holt_data.Rmd")
#true <- data.frame(x = X, y = sapply(X, f, 0, p))
```

@@

```{r}
  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)
  L <- chol(K + sigma_n^2 * I)
  alpha <- forwardsolve(t(L), backsolve(L, obs$y)) # see http://stat.ethz.ch/R-manual/R-patched/library/base/html/chol2inv.html
  k_star <- cov(obs$x, X)
  Y <- t(k_star) %*% alpha
  v <- backsolve(L, k_star)
  Var <- cov(X,X) - t(v) %*% v
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
```

@@

``` {r}
  d <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "direct")
  Ef <- d$Ef
  Cf <- d$Cf
```

@@

``` {r}
s <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "sequential")
ef <- s$Ef
cf <- s$Cf

@@

``` {r}
k <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "sequential")
y_p <- k$Ef
```

@@
